---
title: "Correlation"
subtitle: "Describing the relationship between two numerical variables"
author: "Stat 20 UC Berkeley, Fall 2021"
output:
  xaringan::moon_reader:
    css: stat20-theme.css
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: atelier-forest-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = TRUE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(infer)
library(knitr)
library(xaringanthemer)
library(kableExtra)
library(ggrepel)
source("https://raw.githubusercontent.com/stat-20/stat-20-website/main/stat20-theme.R")
xaringanExtra::use_panelset()
set.seed(401)
```

---
## Announcements
--

- Midterm II
  - Goal: grades and topics published Saturday evening
  - Retake: Sunday 11 am - Monday 11 am
- PS like normal next week
- Lab Tuesday, not Thursday (holiday)


---
class: center, middle, inverse

.adage[Welcome to Part III]

.adage[of Stat 20]

---
## 

.pull-left[
## Part I

**Describing Data**
- Numerical and graphical summaries
- Data fundamentals
  - Data frames
  - Data taxonomy
- Data collection
  - Obs. studies vs experiments
  - Random assignment and causation
- (Probability)
]

--

.pull-right[
## Part II

**Classical Inference**
- Confidence intervals
  - Bootstrapping
- Hypothesis tests
  - Permutation and point nulls
- Mathematical approximations

]

---
## Part III

**Regression**
- Models for multivariate data
  - Simple linear regression
  - Multiple linear regression
  - Logistic regression
- Uses
  - Description
  - Prediction
  - Inference

---
class: center, middle

.adage[Please take out a piece of paper and pencil.]

---
## Example: Poverty and Graduation
--

What is the relationship between _poverty rate_ and _graduation rate_ at the state level?

--

1. Sketch the data frame that can answer this question.
  - Dimensions
  - Unit of observation
  - Column names
  - Range of values

2. Sketch a plot that captures this relationship.
  - How are you applying the grammar of graphics?
  - How would you describe the relationship in words?

--

```{r echo = FALSE}
countdown::countdown(minutes = 2)
```
  
---
## Discuss: Poverty and Graduation

What is the relationship between _poverty rate_ and _graduation rate_ at the state level?


1. Sketch the data frame that can answer this question.
  - Dimensions
  - Unit of observation
  - Column names
  - Range of values

2. Sketch a plot that captures this relationship.
  - How are you applying the grammar of graphics?
  - How would you describe the relationship in words?

```{r echo = FALSE}
countdown::countdown(minutes = 2)
```

---

## The data frame

.pull-left[
1. Sketch the data frame that can answer this question.
  - Dimensions
  - Unit of observation
  - Column names
  - Range of values
]
--
.pull-right[
```{r echo = FALSE}
poverty <- read_delim("figs/poverty.txt")
poverty %>%
  select(State, Graduates, Poverty)
```
]

---

.panelset[
.panel[.panel-name[Question]
2. Sketch a plot that captures this relationship.
  - How are you applying the grammar of graphics?
  - How would you describe the relationship in words?
]
.panel[.panel-name[Grammar]

```{r gg1, eval = FALSE}
ggplot(poverty, aes(x = Poverty, 
                    y = Graduates)) +
  geom_point()
```

- Map `Poverty` to the x-axis.
- Map `Graduates` to the y-axis.
- Use the point geometry.

]
.panel[.panel-name[Plot]
.pull-left-wide[
```{r echo = FALSE, fig.width=10}
p1 <- ggplot(poverty, aes(x = Poverty, 
                    y = Graduates)) +
  geom_point() +
  theme_bw(base_size = 18)
p1
```
]
.pull-right-narrow[
The trend is:

- linear
- moderately strong
- negative
  
> How can we capture this notion in a _statistic_?
  ]
]
]

---
class: center, middle

boardwork

---

.panelset[

.panel[.panel-name[Question]
```{r echo = FALSE, fig.width=12, fig.height=5.5}
library(openintro)
library(ggpubr)
p_scatter <- simulated_scatter %>%
  filter(group %in% c(9:12, 14:17)) %>%
  mutate(group_label = rep(LETTERS[1:8], each = 50)) %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point(size = 2, alpha = 0.8) +
  theme_void(base_size = 18) +
  facet_wrap(~group_label, nrow = 2, scales = "free") + 
  theme(
    panel.border = element_rect(colour = "gray", fill = NA, size = 1),
    strip.background = element_blank()
  )
p_scatter
```

.task[
Which of the following relationships is weakest, as measured by $r$?
]

]

.panel[.panel-name[Responses]
<center>
<iframe src="https://embed.polleverywhere.com/multiple_choice_polls/UsYXLVmOwZcvgXWXwng1U?controls=none&short_poll=true" width="600px" height="450px"></iframe>
</center>
]

.panel[.panel-name[Answer]
```{r echo = FALSE, fig.width=12, fig.height=5.5}
p_scatter + 
  stat_cor(aes(label = paste("r", ..r.., sep = "~`=`~")))
```

**Remember**: r measures only the strength of the _linear_ relationship.
]

]


---
## Poverty and Graduation: Correlation
--

.pull-left[
```{r echo = FALSE, fig.height = 5}
p1
```

```{r cor, eval = FALSE}
poverty %>%
  summarize(r = cor(Graduates, Poverty))
```
]
--
.pull-right[
```{r ref.label = "cor", echo = FALSE}
```

The relationship between poverty and graduation is _linear_, _negative_ and _moderately strong_.

> What other questions can we ask?

]