---
title: "Introduction to Inference"
author: "Stat 20 UC Berkeley, Fall 2021"
output:
  xaringan::moon_reader:
    css: stat20-theme.css
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: atelier-forest-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = TRUE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(knitr)
library(xaringanthemer)
library(kableExtra)
library(emo)
library(palmerpenguins)
library(stat20data)
data(flights)
source("https://raw.githubusercontent.com/stat-20/stat-20-website/main/stat20-theme.R")
xaringanExtra::use_panelset()
```

# Announcements

- Midterm 1 Again 11 am Thurs - 11 am Fri

--

- Evening Study Session
    - 5 - 7 pm Thurs Evans 3rd floor

--

- Lab 5 due _next_ Friday 8 pm
    - Part I Thursday in lab
    - Part II Tuesday in lab

--

- No problem set this week

--

- No lecture Friday


---
class: middle, center, inverse

# Is Yawning Contagious?

---

<div style="position:relative;padding-bottom:56.25%;height:0;overflow:hidden;"> <iframe style="width:100%;height:100%;position:absolute;left:0px;top:0px;overflow:hidden" frameborder="0" type="text/html" src="https://www.dailymotion.com/embed/video/x6tryad" width="100%" height="100%" allowfullscreen > </iframe> </div>

---

## In pairs/trios

.task[
Please record answers to the 6 questions on the front of the handout (the side with **Is Yawning Contagious?** on the front).
]

```{r}
countdown::countdown(3)
```

--

.task[
Please enter your answers at `pollev.com/andrewpbray088`
]

