---
title: "Lab 2: Sampling"
output: github_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set()
library(tidyverse)
```

The quintessential activity that Reed alumni do upon returning to campus is head up to the thesis tower to check to see if anyone has checked out their senior thesis. They're wondering: what is the chance that someone was interested in my work? To answer that question and learn more generally about the way in which senior theses are checked out, you will collect data on this phenomenon.

## Recording Data

Your group will enter their data in a spreadsheet found here: [https://bit.ly/38shltt](https://bit.ly/38shltt).

Each group will have their data in a different sheet of this google doc. Click the plus in the lower-left corner to add a new sheet, rename by concatenating your first names with `_`. To make life easier down the line, please follow the division name guidelines on the first example sheet.

1. What is your guess of the total number of theses in the tower? Enter your guess here as well in the google sheet.

Add the following row of data to your google sheet. 

```{r echo = FALSE}
library(knitr)
data.frame(barcode = 12345678,
           author_last_name = "tester",
           year = 999,
           checkouts = 1,
           division = "MNS",
           n_pages = 100) %>%
  kable()
```

Next, try running the following code to test if you're able to bring the data set into R:

```{r eval = FALSE, echo = TRUE}
options(httr_oob_default = TRUE)
library(tidyverse)
library(googlesheets4)
url <- "https://docs.google.com/spreadsheets/d/14DG_oUyg4bHFkQZsIgWVF9A042lw4ft9uWVAI4yTqyY/edit?usp=sharing"
theses <- read_sheet(ss = url,
           sheet = "andrew",
           skip = 11)
```

Be sure to swap out `"andrew"` with the name of your group's sheet.

## Sampling considerations

Your objective is to learn as much about senior theses checkout patterns as you can with 3 people and only about an hour. You need to record six items of information on each thesis; `barcode` number, `author_last_name`,  `year` of publication, number of `checkouts`, and `division` (all departments are grouped into divisions, found at http://www.reed.edu/academics.html). 

Since your resources are limited, sampling is an essential technique here. With your group, discuss the following considerations to help guide your sampling. Keep in mind the example of literary digest - more data is worthless if it's not representative - and spend 15 - 30 minutes on this section. Also, you may want to collect preliminary data, i.e. try out your method on a very small population to assess if it is working as you intended and how long it takes. Please detail your sampling strategy below.

2. What organizational scheme do the theses follow in the tower?

3. Please do your best to estimate the total number of theses in the tower (also record this in your google sheet).

4. Please provide a detailed plan of how you will collect a sample of data that is representative and efficient? You might consider using a simple random sample, stratified sampling, or cluster sampling. Your plan should be explicit enough that someone else could follow in your footsteps next semester and know how to collect a sample in an identical manner.


## Data Collection and Reflections

Now is your opportunity to draw a sample of theses and record the variables of interest. Please wrap things up in the tower room at the end of class: even if you'd like to have a larger sample, a small thoughtful data set can still be very valuable. Before 6 pm this evening, your group is responsible for entering that data into your google sheet.

After you have entered in all of your data, please answer the following questions. Don't forget to load your data set into your Rmd! You can find the code for that in the chunk of code at the top of this lab. To keep things clean for that chunk, use the following R chunk options: ````{r echo = FALSE, eval = TRUE, message = FALSE}`.

5. Did you encounter any unexpected challenges when collecting this data? How did you resolve them?

6. Create a plot to visualize the relationship between the number of checkouts and the year of publication. Describe the structure that you find. How many checkouts can you expect in your thesis if you come back for your 10th reunion? Your 20th?

7. Create a plot that addresses the question: "Is there a trend in length of theses over time?". Answer that question also in words.

8. Create a plot that addresses the question: "Are theses from the different divisions checked out at different rates?". Answer that question also in words.

9. Combine the variables from plots 6 and 8 into a single plot that shows the trends in checkouts over time, separated out for each division. Describe any structure that you see.
