---
title: "Lab 1: Data Visualization"
output: github_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(dplyr)
library(ggplot2)
library(readr)
library(oilabs)
data(survey141)
```

## Setup

This lab will focus on generating plots using the `ggplot2` package, as well as 
practicing data subsetting with the `filter()` command from the `dplyr` package, both of which live in the `tidyverse` package.

To get started, load the following packages and data set.

```{r load-packages, message=FALSE}
library(tidyverse)
library(oilabs)
data(survey141)
```

This data comes from a class survey that Math 141 students took two years ago, however the data frame 
we'll be using has been slightly altered to make generating plots easier. The column names 
of the data set have been changed as well. The help file contains the original questions. 

```{r}
?survey141
```

## Plot Types

The `ggplot` command is used to create a base layer for a plot, with `aes()` used to specify
which visual cues to connect to variables in a data frame. After creating this base layer, a
geometry can be applied which will render a specific type of plot. 

```{r eval = TRUE}
d <- ggplot(survey141, aes(x = hot_dog, fill = dog_pants))
d + geom_bar()
```

Above a base layer for the survey data is created with `hot_dog` mapped to the x-axis and `dog_pants` mapped to fill (a color 
based visual cue). This base layer is assigned to `d` and then a bar plot layer is added to it via `geom_bar()`.

The following geoms are some of the most commonly used:

    - `geom_bar()`:         for a categorical variables
    - `geom_histogram()`:   for a numerical variable
    - `geom_density()`:     for a numerical variable
    - `geom_boxplot()`:     for a numerical variable and possibly a categorical variable
    - `geom_point()`:       for 2 numerical variables
    - `geom_smooth()`:      for 2 numerical variables
    
Multiple geometries may be layered on top of each other, as is often the case with 
`geom_smooth()` and `geom_point()` to create scatter plots with trend lines.
    
1. Create a density plot for the number of colleges applied to. Please describe this
distribution in terms of its shape, center, and spread.

```{r echo = FALSE, eval = FALSE, message = FALSE, warning = FALSE}
ggplot(survey141, aes(x = college_app)) + geom_density()

We haven't yet formally covered the measures of shape, center, and spread, so assess their descriptions holistically.
```




## Filtering Data

Often there will be entries in data sets that you'll want to exclude from a particular analysis. 
This is when the `filter()` command will be useful. `filter()` will take a 
data set and find all rows that return true under set conditions (or filters). 

The following example has the `survey141` data being filtered to only include 
rows which have "Tea" as the response for the `coffee_tea` variable. This subset 
of the original survey data is then assigned to a new data frame called `tea_drinkers`.

```{r}
tea_drinkers <- filter(survey141, coffee_tea == "Tea")
```

The following uses the piping operator to apply a filter to `survey141`, filtering out only
those rows which have 0 in the `marijuana` variable, **and** a value greater than 0 in the
`alcohol` variable. 

```{r}
alcohol_only <- survey141 %>% 
  filter(alcohol > 0 & marijuana == 0)
```

Pulling up the help file will give a list of useful filter functions and more examples
of the function in use.

2.  Filter out those who answered PC or Mac to the computer of choice question. 
Create a side-by-side box plot comparing the first kiss age between the groups.
Do you notice any differences between the two groups? If so, what are they?
(Note: if the box plot does not render correctly, try switching the `x` and `y` assignments.)

```{r echo = FALSE, eval = FALSE, message = FALSE, warning = FALSE}
filter(survey141, computer == "PC" | computer == "Mac") %>%
  ggplot(aes(x = computer, y = first_kiss)) + 
  geom_boxplot()

There appear to be no difference in the medians of first kiss age between the two groups, althrough there is evidence that the Mac users have a bit longer left tail. Each group has one low outlier.
```

## Settings, Labels and Limits

Beyond aesthetics and geometries, `ggplot2` also has settings and labels for customizing 
visualizations.

Settings are options one can apply to a plot, but unlike aesthetics, they are not directly linked to a variable.

```{r eval = TRUE}
ggplot(survey141, aes(x = hot_dog, fill = dog_pants)) +
  geom_bar(position = "fill")
```

In the above example, the setting `position = "fill"` is used in `geom_bar()` so that count is no longer 
displayed, but instead proportion. 

3.  Using `position = "jitter"` create a scatter plot relating `alcohol` and `marijuana`. 

```{r echo = FALSE, eval = FALSE, message = FALSE, warning = FALSE}
p <- ggplot(survey141, aes(x = alcohol, y = marijuana)) + 
  geom_point(position = "jitter")
p
```

Beyond geoms, additional layers such as `labs()` and `xlim()` can be added to a plot, used to 
add labels/titles and axis limits respectively. These layers are added on just like a geom 
and can help add clarity or focus to a specific part of a plot.

4. Give more descriptive axis labels and specify limits for the previous plot to remove more 
extreme cases. Describe any relationship or lack thereof. How do these last two plots compare to 
your personal conception of how these variables are related?

```{r echo = FALSE, eval = FALSE, message = FALSE, warning = FALSE}
p + xlim(-1, 5) + 
  ylim(-1, 5) + 
  labs(x = "something", y = "something else")

Answers will vary, though there is a weak to moderate association between these two variables in this data set.
```

5.  Create 2-3 more visualizations from the survey data that are of interest to you.
Have at least one visualization come from a filtered subset of the data and have at
least one more map 3 or more variables. Describe any trends/patterns discovered.

```{r echo = FALSE}
# After some time working on this have people share what they've found

Answers will vary.
```

6.  Recreate exactly the following plot. The help file for `geom_smooth()` will be useful to get the best fit lines looking appropriate.

```{r echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
ggplot(survey141, aes(x = economic, y = social, color = year)) + 
  geom_point(position = "jitter") + 
  geom_smooth(method = "lm", se = FALSE) + 
  xlim(0,10) +
  ylim(0,10) + 
  labs(title = "Political Views and Year", 
       subtitle = "1 - Extremely Liberal, 10 - Extremely Conservative", 
       x = "Economic Views",
       y = "Social Views",
       color = "Year")
```


## More Practice

[Fivethirtyeight.com](https://fivethirtyeight.com/) is a news website that focuses on data journalism: nearly all of their stories are inspired by a data set. The editors of the website make that data available on github for all to  explore, and several fine members of R community put it into an R package so that we can have easy access to it. Load in the package and view the names of the data sets with the following.

```{r eval=FALSE}
library(fivethirtyeight)
data(package = "fivethirtyeight")
```

You can learn more about the listed data sets by pulling up with helpfile by running a command like `?bob_ross`. If that seems a data set of interest to you, you can load it into R and view it with:

```{r eval = FALSE}
data("bob_ross")
View(bob_ross)
```

7. Select a data set (not `bob_ross`)  and produce three plots. There are countless plots that you *could* generate, but aim to make plots that reveal interesting structure in the data and are throughfully constructed. Be sure to add useful axis labels and a title plot. Provide a sentence or two for each plot describing what it reveals about the subject matter. Hint: usually data sets with a mix of categorical and numberical data are easier to visualize.

```{r eval = FALSE, echo = FALSE}
Answers will vary.
```


8. Only after you have produced these three plots, go ahead and read the story that this data set inspired. Those can be found by clicking into the appropriate folder at [this website](https://github.com/fivethirtyeight/data) and following the link. Try to reproduce one of the plots featured in the story. Add a sentence or two describing what the plot shows in terms of the data and how your plot differs from the one published in the story. If you aren't able to reproduce any of the plots, describe precisely the hurdles that you ran into. Be sure to include in your answer a link to the original plot in the story. You can get this by right-clicking on the image, selecting "copy image address", then pasting into a markdown section of your document with the syntax `[the original image can be found here](here-is-the-pasted-url)`.

```{r eval = FALSE, echo = FALSE}
Answers will vary.
```