---
title: "Lab 1: Introduction to R and RStudio"
output: github_document
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(message = FALSE, eval = FALSE)
library(tidyverse)
library(oilabs)
```


## Dr. Arbuthnot's Christening Records

To get you started, run the following command to load the data.

```{r load-abrbuthnot-data, eval=TRUE}
data(arbuthnot)
```

The Arbuthnot data set refers to Dr. John Arbuthnot, an 18<sup>th</sup> century 
physician, writer, and mathematician. He was interested in the ratio of newborn
boys to newborn girls, so he gathered the baptism records for children born in
London for every year from 1629 to 1710. We can view the data by 
typing its name into the console.

```{r view-data, eval = FALSE}
arbuthnot
```

An alternate way to look at the data is to use the data viewer. Click on
the name `arbuthnot` in the *Environment* pane (upper right window) that lists 
the objects in your workspace. This will bring up an alternative display of the 
data set in the *Data Viewer* (upper left window). You can close the data viewer
by clicking on the `x` in the upper lefthand corner.

What you should see are four columns of numbers, each row representing a 
different year: the first entry in each row is simply the row number (an index 
we can use to access the data from individual years if we want), the second is 
the year, and the third and fourth are the numbers of boys and girls baptized 
that year, respectively. Use the scrollbar on the right side of the console 
window to examine the complete data set.

Note that the row numbers in the first column are not part of Arbuthnot's data. 
R adds them as part of its printout to help you make visual comparisons. You can
think of them as the index that you see on the left side of a spreadsheet. In 
fact, the comparison to a spreadsheet will generally be helpful. R has stored 
Arbuthnot's data in a kind of spreadsheet or table called a *data frame*.

You can see the dimensions of this data frame as well as the names of the variables and the first few observations by typing:

```{r glimpse-data, eval = FALSE}
glimpse(arbuthnot)
```

This command should output the following

```{r glimpse-data-result, echo=FALSE, eval=TRUE}
glimpse(arbuthnot)
```

We can see that there are 82 observations and 3 variables in this dataset. The variable names are `year`, `boys`, and `girls`. At this point, you might notice 
that many of the commands in R look a  lot like functions from math class; that 
is, invoking R commands means supplying a function with some number of arguments. 
The `glimpse` command, for example, took a single argument, the name of a data frame. 

## Some Exploration

Let's start to examine the data a little more closely. We can access the data in
a single column of a data frame separately using a command like

```{r view-boys}
select(arbuthnot, boys)
```

This command will only show the number of boys baptized each year. This can be read
as "select the boys columns out of the arbuthnot data frame."

1. What command would you use to extract just the counts of girls baptized? Try
    it!


### Data visualization

R has some powerful functions for making graphics. We can create a simple plot 
of the number of girls baptized per year with the command

```{r plot-girls-vs-year}
ggplot(arbuthnot, aes(x = year, y = girls)) +
         geom_point()
```

In the `ggplot()` function we specify which columns of the data frame we wish to visualize,
as well as our desired geometry (or geom). If we wanted to connect the data points with 
lines, we could instead use the `geom_line()`.

```{r plot-girls-vs-year-line}
ggplot(arbuthnot, aes(x = year, y = girls)) +
         geom_line()
```

2. Is there an apparent trend in the number of girls baptized over the years? 
How would you describe it? (To ensure that your lab report is comprehensive,
be sure to include the code needed to make the plot as well as your written
interpretation in markdown.)

### Adding a new variable to the data frame

It will be useful for future analysis if we had access to additional column of information:
the total number of births in every year. We can add this with the `mutate()` function.

```{r calc-total-bapt-vars-save}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
```

The `%>%` operator is called the **piping** 
operator. It takes the output of the previous expression and pipes it into 
the first argument of the function in the following one. 
To continue our analogy with mathematical functions, `x %>% f(y)` is 
equivalent to `f(x, y)`.

**A note on piping:** Note that we can read these three lines of code as the following: 

*"Take the `arbuthnot` dataset and **pipe** it into the `mutate` function. 
Mutate the `arbuthnot` data set by creating a new variable called `total` that is the sum of the variables
called `boys` and `girls`. Then assign the resulting dataset to the object
called `arbuthnot`, i.e. overwrite the old `arbuthnot` dataset with the new one
containing the new variable."*

This is equivalent to going through each row and adding up the `boys` 
and `girls` counts for that year and recording that value in a new column called
`total`.

You'll see that there is now a new column called `total` that has been tacked on
to the data frame. The special symbol `<-` performs an *assignment*, taking the 
output of one line of code and saving it into an object in your workspace. In 
this case, you already have an object called `arbuthnot`, so this command updates
that data set with the new mutated column.

We can make a plot of the total number of baptisms per year with the command

```{r plot-total-vs-year}
ggplot(arbuthnot, aes(x = year, y = total)) +
         geom_line()
```

Similarly to how we computed the total number of births, we can compute the ratio 
of the number of boys to the number of girls baptized in 1629 with

```{r calc-prop-boys-to-girls-numbers}
5218 / 4683
```

or we can act on the complete columns with the expression

```{r calc-prop-boys-to-girls-vars}
arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)
```

We can also compute the proportion of newborns that are boys in 1629

```{r calc-prop-boys-numbers}
5218 / (5218 + 4683)
```

or this may also be computed for all years simultaneously and append it to the dataset:

```{r calc-prop-boys-vars}
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)
```

Note that we are using the new `total` variable we created earlier in our calculations.

3. Now, generate a plot of the proportion of boys born over time. What do you see?

Finally, in addition to simple mathematical operators like subtraction and 
division, you can ask R to make comparisons like greater than, `>`, less than,
`<`, and equality, `==`. For example, we can ask if boys outnumber girls in each 
year with the expression

```{r boys-more-than-girls}
arbuthnot <- arbuthnot %>%
  mutate(more_boys = boys > girls)
```

This command add a new variable to the `arbuthnot` dataframe containing the values
of either `TRUE` if that year had more boys than girls, or `FALSE` if that year 
did not (the answer may surprise you). This variable contains a different kind of 
data than we have encountered so far. All other columns in the `arbuthnot` data 
frame have values that are numerical (the year, the number of boys and girls). Here, 
we've asked R to create *logical* data, data where the values are either `TRUE` 
or `FALSE`. In general, data analysis will involve many different kinds of data 
types, and one reason for using R is that it is able to represent and compute 
with many of them.

* * *

## More Practice

In the previous few pages, you recreated some of the displays and preliminary 
analysis of Arbuthnot's baptism data. Your assignment involves repeating these 
steps, but for present day birth records in the United States. Load the 
present day data with the following command.

```{r load-present-data}
data(present)
```

The data are stored in a data frame called `present`.

4.  What years are included in this data set? What are the dimensions of the 
    data frame? What are the variable (column) names?

5.  How do these counts compare to Arbuthnot's? Are they of a similar magnitude?

6.  Make a plot that displays the proportion of boys born over time. What do you see? 
    Does Arbuthnot's observation about boys being born in greater proportion than girls 
    hold up in the U.S.? Include the plot in your response. *Hint:* You should be
    able to reuse your code from Ex 3 above, just replace the dataframe name.

7.  In what year did we see the most total number of births in the U.S.? *Hint:*
    First calculate the totals and save it as a new variable. Then, sort your
    dataset in descending order based on the total column. You can do this
    interactively in the data viewer by clicking on the arrows next to the 
    variable names. To include the sorted result in your report you will need 
    to use two new functions: `arrange()` (for sorting). We can arrange the data 
    in a descending order with another function: `desc()` (for descending order). To
    learn the syntax for this, bring up the help file for `arrange()` by typing `?arrange`
    at the console.

These data come from reports by the Centers for Disease Control. You can learn more about them
by bringing up the help file using the command `?present`.


<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>
